The Globular Cluster System of the Auriga Simulations

### 2019-06-11 AurigaGCS.tex
Start over w/ introduction. Sorta satisfied w/ first three paragraphs now

\section{Introduction}
Globular clusters (GC)s are omnipresent, bright (with ${<}M^0_V{>} = -7.1 \pm 0.4$
that is nearly independent of the host galaxy size and environment,
\citealt{1991ARA&A..29..543H}), old ($\gtrsim$10~Gyr, \citealt{2005A&A...439..997P,
2005AJ....130.1315S, 2009ApJ...694.1498M, 2013ApJ...775..134V}), and amongst the
simplest stellar configurations in the Universe. The typical luminosity gives
a characteristic mass of $1.4\cdot 10^5$~M\Sun \, (assuming a mass to light ratio
$M/L_V \simeq 2$), and GCs have typical half-mass radii $r_h$ of a few parsec.
GCs within the Milky Way can thus be studied in great detail (at the level of 
individual stars), whereas integrated quantities can be explored with reasonable 
certainty for GCs within - and even beyond \citep[e.g.][]{1991ARA&A..29..543H} -
the Local Group.

The extensive literature on GCs is summarised in several 
books and review articles \citep[e.g.][]{1991ARA&A..29..543H, Harris2001, 2004Natur.427...31W, 
2006ARA&A..44..193B, 2012A&ARv..20...50G, 2014CQGra..31x4006K, 2018RSPSA.47470616F},
and recent analyses of the Gaia data shed new light on the chemical- and structural 
properties, and on the phase-space distribution of the MW GC system \citep[e.g.][]{
2018A&A...616A..12G, 2018ApJ...859L..13B, 2019A&A...621A..56P, 2019MNRAS.484.2832V}.
A clear picture of the formation GC systems has yet to emerge, despite decades 
of studies and the tremendous advances the last couple of years. It is clear, 
however, that a fundamental understanding of GC formation and evolution promises 
exciting insight into the detailed histories of their host galaxies due to their
old ages and several correlations between GC systems and their parent galaxies.

% There are several correlations between globular cluster systems and their
% host galaxy. For example, the total mass of globular cluster systems
% correlates with the halo mass \citep[e.g.][]{2010MNRAS.406.1967G}, the
% the total number of GCs increases linearly with increasing virial mass
% \citep[e.g.][]{2019arXiv190100900B}, and THIRD EXAMPLE.

The literature offers a wealth of formation channels, most of which can be broadly
classified as ex situ, or as in situ scenarios. In the former picture, GCs form
in smaller, more metal-poor progenitors that are accreted in subsequent mergers,
as can naturally be expected in the framework of hierarchical galaxy formation
\citep[e.g.][]{1978ApJ...225..357S, 1984ApJ...277..470P}, and Rosenblatt et al. 1988;
Cen 2001; Bromm \& Clarke 2002; Diemand et al. 2005; Moore et al. 2006; Boley et al. 2009;
Griffen et al. 2010; Trenti et al. 2015; Kimm et al. 2016; Ricotti et al. 2016;
Boylan-Kolchin 2017; Creasey et al. 2019. As for in situ scenarios,
\citet{1985ApJ...298...18F} propose that thermal instabilities in the collapsing
proto-galaxy may lead to GC formation, \citet{1992ApJ...384...50A, 2010ApJ...718.1266M,
2018MNRAS.480.2343C} suggest that GC formation is triggered by (major) gas-rich
galaxy-galaxy mergers and/or interactions, and \citet{2005ApJ...623..650K} envision
GCs to form in the cores of supergiant molecular clouds located within gaseous disks
of (high-redshift) galaxies.

Another broad distinction between formation mechanisms are models that rely on
special conditions in the early Universe, or those that test the hypothesis that
GC formation at high redshift is governed by the same physical processes of young
massive clusters \citep[see][for a review]{2010ARA&A..48..431P} in the Local
Universe, where differences arise as the result of nearly a Hubble time of
(dynamical) evolution. This (latter) scenario is extensively tested in the recent
E-MOSAICS project \citep{2018MNRAS.475.4309P, 2019MNRAS.486.3134K}, which
comprises a suite of cosmological zoom simulations of MW-like galaxies. As for
the former category, \citet{2017Ap&SS.362..183R} investigate what fraction of
(multiple) stellar populations ends up in bound star clusters after their
formation in gravitationally unstable expanding supershells, which arise from
hypernova explosions of massive pop. III stars. \citet{2019arXiv190408941C}
speculate in a recent study that GCs may form in supersonically induced gasseous
objects at high redshifts, and \citet{2019arXiv190508951M} propose that high speed
collisions leave cooling gas without their DM subhaloes which could form GC
populations that can reproduce the observed $N_{\text{GC}}-M_{\text{vir}}$
correlation \citep[e.g.][]{2019arXiv190100900B} with bimodal metallicity
distributions and ages consistent with observations.

\citep{2019arXiv190505199C}
``The  mean  and  dispersion  of  GC system MDFs have also been found to scale 
with the mass of the host galaxy over a large range in galaxy mass (Peng et al.
2006). Qualitatively, many galaxies have clearly multi-modal MDF (e.g., Gebhardt
\& Kissler-Patig 1999; Peng et al. 2006; Brodie \&  Strader  2006;  Brodie  et  
al.  2012),  but  the  most massive galaxies typically show broad, single-peaked 
distributions (Harris et al. 2014, 2016, 2017a)''

One important constraint on plausible formation mechanisms comes from observed
bimodality in the colour distribution of most galaxies, believed to be indicative
of bimodality in the underlying metallicity distribution of GC systems. GCs in
the MW can indeed be split up into a metal-rich (red), and a metal-poor (blue)
(sub)population. The former group is associated to the galactic disk because it
is a flattened, rotating population found at radii smaller than solar. The latter,
on the other hand, is more radially extended, more populated (with three times
more blue than red GCs), and may be $\leq 1.5$~Gyr older \citep{2005AJ....130..116D}.
The exact dividing metallicity slightly differs between authors, e.g.
[Fe/H]~$=-0.8$ \citep[according to][]{1985ApJ...293..424Z}, or
[Fe/H]~$=-1.0$ \citep[according to][]{Harris2001}, and may also differ between
galaxies \citep[e.g.][]{1985ApJ...293..424Z, 1999AJ....118.1526G, 2001AJ....121.2974L,
2006ApJ...639...95P}. In contrast, it has long been debated whether the Andromeda
GCS metallicity distribution is uni- bi- or multimodal, and \citet{2016ApJ...824...42C}
argue that the data favours three groups (split at [Fe/H] $= -0.4$ and $-1.5$)
after more accurate age-measurements improved classification of star clusters
as globular (or otherwise). Moreover, the Andromeda GCS has $2-3$ times as many
GCs at large radii, believed to be indicative of a more rich merger history.

The latter is supported by observed similarities between the spatial distribution
and chemical signature of the `blue' GCs and the Galactic stellar halo
\citep{2008A&ARv..15..145H}. Moreover, relative age-estimates indicate that the
metal-poor subpopulation may be $1.5$~Gyr younger than the metal-rich counterpart
\citep{2005AJ....130..116D}, providing further support that the GC population indeed
consists of two distinct classes.

People postulate that the frequent mergers that occur during hierarchical build-up
of galaxies naturally produce (sub)populations of GCs from enriched gas on tight
orbits, if the mass ratios and gas fractions of the mergers are sufficiently high.

Boylan-Kolchin: the blue population could form in high density regions along the
cosmic filament before or during the collapse of the proto-galaxy itself.
This formation scenario would explain the presence of the oldest GCs with
primordial metallicities in the (outer) haloes of galaxies as the GCs are
accreted over time. Moreover, the tidal fields experienced by GCs at (outer) halo
orbits is significantly lower than by GCs in the inner galaxy, thus such GCs
are expected to be long-lived.

An alternative paradigm is that GC formation does not rely on special conditions
at high redshifts. Rather, the same physics that governs the formation of young
massive clusters (YMCs) in the Local Group can be applied to the high redshift
Universe. The observed differences between YMCs and GCs, in this scenario, arise
naturally as the result intrinsic GC evolution within galactic tidal fields over
a Hubble time. This interesting hypothesis is tested numerically by the E-MOSAICS
people.

\citet{2010MNRAS.404.1203F} collected age-metallicity observations from the
literature and found one track with constant age of 12.8~Gyr, and one track
with branches towards younger ages. These branches are dominated by GCs associated
with the Sagittarius, Canis Major and other accreted dwarf galaxies, ultimately
indicating that 27-47 GCs (about one 25\% of the MW GCS) have been accreted
from six to eight dwarf galaxies. \citet{2019MNRAS.486.3134K, 2019MNRAS.486.3180K}
show that GC simulated age-metallicity space of the E-MOSAICS suite indeed allows
for a reconstruction of the merger history of the Milky Way-like galaxies, and
conclude that 40\% of the MW GCS has been accreted from galaxies with a stellar
mass range of $2 \cdot 10^{7} - 2 \cdot 10^{9}$ M\Sun, and $6\pm1$ are core remnants
of accreted dwarfs.

% \begin{itemize}
%     \item \citet{2005MNRAS.364..367D}: ``The radial profile of the stellar halo
%     and metal-poor globular clusters of the Milky Way suggest that these components
%     formed in rare early peaks above $2.5 \sigma$ at redshift above 10. ''
%
%     \item \citet{2017MNRAS.465.3622R}: ``GCs among oldest astrophysical objects.
%     GCs form in the early Universe in highest density peaks
%     \citep[e.g.][]{2005MNRAS.364..367D, 2009ApJ...706L.192B}''
%
%     \item ``Hence, they witness most of the formation and evolution processes
%     of galaxies, and can be used to probe them'' \citep{2006ARA&A..44..193B}
% \end{itemize}

\begin{itemize}
    \item Special conditions in early Universe?
    \item Formation in collapsing proto-galaxies (in-situ-ish?)
    \item Formation as a result of (wet) mergers?
    \item Formation in satellites that are later accreted?
    \item No special conditions at high z, but connection between low-z YMC-
        and high-z GC-formation where differences could result from Hubble time
        of evolution?
    \item This means that in order to test formation models against observations,
        one not only needs needs to model GC formation, but also include the full
        evolution over their lifetime in a cosmological tidal field.
\end{itemize}

\todo[inline]{Paragraph: ingredients for numerical studies}
\begin{itemize}
    \item High resolution in the ISM
    \item Full cosmological evolution
\end{itemize}

% \todo[inline]{Paragraph: scientific context / state of the field: briefly summarise
%     previous work and references work of other groups}
% \begin{itemize}
% \item Origin of the Milky Way globular clusters \citep{2017MNRAS.465.3622R}
% \item GCs in FIRE \citep{2018MNRAS.474.4232K}
% \item EMOSAICS project \citep{2018MNRAS.475.4309P}
% \item GC-DMhalo connection \citep{2017MNRAS.472.3120B}
% \item Origin of GC bimodality? \citep{2018MNRAS.479..200F}
% \item \textit{GAIA} DR2: GC kinematics \citep{2018A&A...616A..12G}
% \item Dating GC Tidal Disruption \citep{2018ApJ...859L..13B}
% \item GC in N-body simulation \citep{2018ApJ...861...69C}
% \item (Role of GC mass evolution on stream properties \citep{2018MNRAS.474.2479B})
% \item GC formation from dwarfs to giants \citep{2018MNRAS.480.2343C}
% \item (GC contribution to EOR \citep{2018MNRAS.479..332B})
% \item Early Universe supermassive star / GC formation \citep{2018MNRAS.478.2461G}
% \item GC formation in cold filaments \citep{2018ApJ...861..148M}
% \item GC formation in high-redshift dwarf galaxies \citep{2018MNRAS.477..480Z}
% \item GCs in MW outer region \citep{2017arXiv170804542P}
% \item Impact of the Cutoff of the Cluster Initial Mass Function \citep{2018arXiv181001888C}
% \item Metallicity gradients in the globular cluster systems of early-type galaxies: in situ and accreted components \citep{2018MNRAS.479.4760F}
% \item Globular clusters in M31, Local Group, and external galaxies \citep{2016IAUS..317..120L}
% \item Globular Clusters Formed within Dark Halos I: present-day abundance, distribution and kinematics \citep{2019MNRAS.482..219C}
% \item (The mass of the Milky Way from satellite dynamics \citep{2018arXiv180810456C})
% \item Globular cluster formation and evolution in the context of cosmological galaxy assembly: open questions \citep{2018RSPSA.47470616F}
% \item The kinematics of globular clusters systems in the outer halos of the Aquarius simulations \citep{2016A&A...592A..55V}
% \item Star Cluster Formation in Cosmological Simulations \citep{2017ApJ...834...69L, 2018ApJ...861..107L, 2018arXiv181011036L}
% \item (A systematic analysis of star cluster disruption by tidal shocks - I. Controlled N-body simulations and a new theoretical model \citep{2018arXiv181200014W})
% \item Spatial mixing of binary stars in multiple-population globular clusters \citep{2018MNRAS.tmp.3147H}
% \item Star Clusters Across Cosmic Time \citep{2018arXiv181201615K}
% \item Kinematics of Subclusters in Star Cluster Complexes: Imprint of their Parental Molecular Clouds \citep{2018arXiv181201858F}
% \item Investigating the population of Galactic star formation regions and star clusters within a Wide-Fast-Deep Coverage of the Galactic Plane  \citep{2018arXiv181203025P}
% \item Globular cluster number density profiles using \textit{Gaia} DR2 \citep{2019arXiv190108072D}
% \item Clustering Clusters: unsupervised learning on GC observations (of log $M$,
%     log $\sigma_0$, log $R_e$ , [Fe/H], log $|Z|$) to determine number of groups
%     $k=2$ (disk/halo) or $k=3$ (disk/inner-halo/outer-halo): the optimal number
%     of groups given the data is $k=3$. \citep{2019arXiv190105354P}.
% \item There is a well-known hypothesis that Open Stars Clusters (OSCs) could form as a result
%     of GCs crossing the disk. New kinematical data does not support
%     this claim for two particular OSC-GC pairs, and does seem consistent
%     with this scenario for six OSC-GC pairs \citep{2019arXiv190106481B}. (Note
%     that GC crossing the disk could also trigger formation of multiple OSCs!)
% \item DM halo mass can be inferred from number of globular cluster in galaxy
%     \citep{2019arXiv190100900B}
% \item GCs in M31 outer halo \citep{2019arXiv190111229W}
% % TODO: last checked arXiv Mar 2, 2019
% \end{itemize}

% \todo[inline]{Paragraph, narrowing open questions in GC research down to formation,
%     possibly ``Bimodality suggests two formation mechanisms''. In-situ
%     formation vs. accretion in hierarchical build-up of galaxies naturally produces
%     two populations of globular cluster.}
% \begin{itemize}
%     \item ``colour bimodality, blue and red clusters \citep[e.g.][]{1985ApJ...293..424Z, 1999AJ....118.1526G, 2001AJ....121.2974L, 2006ApJ...639...95P}
%     \item ``blue metal-poor (with distribution peaking at [Fe/H] $\approx -1.5$ for the Milky Way), no sign of rotation as a population (..) more metal-rich (peak at [Fe/H] $\approx -0.5$ in the Milky  Way) more spatially concentrated and rotating with the galaxy.'' \citep{1996AJ....112.1487H}
%     \item ``Blue clusters from in early Universe in galaxies that merge later. In (wet) merger process starbursts generate red population \cite{1992ApJ...384...50A, 1987nngp.proc...18S}''
%     \item ``\citet{1997AJ....113.1652F} propose instead that blue globulars form when the protogalaxy itself collapses, in a metal-poor and turbulence media. The red population would form later, once the galactic disc has settled. The formation of globular clusters would then be a multiphase process, with the first phase being interrupted possibly by cosmic reionization \citep{2002MNRAS.333..383B}.''
%     \item ``\citet{2005ApJ...623..650K, 2014ApJ...796...10L} advocate
% that major mergers are at the origin of both sub-populations: blue
% clusters form during early mergers (z > 4) while the red ones appear
% in mergers at lower redshifts (even after z = 1). Although, this
% scenario, combined with star formation enhancement in mergers,
% seems appropriate in dense galactic environment leading to the
% assembly of massive elliptical galaxies, like in the Virgo Cluster
% as tested by \citet{2014ApJ...796...10L}, it does not apply to Milky Way-
% like systems where no recent major merger took place (Wyse 2001;
% Deason et al. 2013; Ruchti et al. 2014, 2015).
% \item ``\citet{1998ApJ...501..554C} argue that red clusters form in situ while the blue ones are accreted, either via merging satellite galaxies, or by tidal capture of the clusters themselves \citep[see also][]{2013ApJ...762...39T}.''
% \end{itemize}

\todo[inline]{Paragraph: narrow down to this work}
The star formation model implemented in the Auriga simulations is capable of
producing realistic Milky Way-like (spiral disk) galaxies at redshift zero. Therefore
the question naturally arises whether or not the Auriga simulations also produce
a subset of star particles (GC candidates) which is consistent with the GC system
of the Milky Way and/or the Andromeda galaxy (M31). The typical mass scale of
globular clusters is comparable to the numerical (mass) resolution of cosmological
zoom simulations. The detailed physics of GC formation in the Universe appears in
observations as the combined effect of the $10^{3-6}$ M\Sun, compared to a mass
resolution of $10^{3-5}$ M\Sun in the Auriga simulations. Globular clusters can
therefore serve as an ultimate test to the star formation model that is implemented
in the numerical simulations. Moreover, cosmological zoom simulations provide a
detailed (formation and merger) history of the simulated galaxies. This is
important because theoretical paradigms for globular cluster formation in the
literature know two distinct classes of GCs that are separated by their exact
formation sites: an in situ versus an accreted population. Cosmological zoom
simulations uniquely allow for an investigation into globular cluster formation
with a particular focus on the properties of the in situ versus accreted subpopulations.

\todo[inline]{Paragraph: Paper outline}
We summarise the relevant characteristics of the Auriga simulations in
section~\ref{sec:auriga}, followed by a summary of the observations of the
Milky Way (MW) globular cluster system (GCS) in section~\ref{sec:observations}
that we use to compare our simulations to in section~\ref{sec:results}. We
discuss our findings in section~\ref{sec:discussion} to come to our conclusions
in section~\ref{sec:conclusions}.

>>>>

\section{Introduction}
Globular clusters (GC)s are old, bright, ubiquitous, and various properties of
GC systems show correlations with their host galaxies. GC systems are believed to
retain information about the galactic (gas) conditions at times of formation,
thus, could offer unique insight into the chemodynamical evolution of their 
parent galaxies, if the formation and evolution of GCs- and GC systems themselves 
is adequately understood. However, despite decades of research (the extensive 
literature on GCs is summarised in several books and review articles, e.g.~\citealt{
1991ARA&A..29..543H, Harris2001, 2004Natur.427...31W, 2006ARA&A..44..193B, 
2012A&ARv..20...50G, 2014CQGra..31x4006K, 2018RSPSA.47470616F}), consensus on an 
exhaustive picture of the formation of GCs is yet to be reached.

The literature offers a wealth of formation scenarios. \citet{1968ApJ...154..891P}
and \citet{1984ApJ...277..470P} argue that GCs form as the earliest bound
structures in the Universe (i.e. prior to formation of the main galaxy), 
noting that the Jeans length and mass shortly after recombination is consistent
with typical GC masses and sizes. Alternative models suggest that GCs form during
formation of proto-galaxies, for example as a result of thermal instabilities
in hot gas-rich haloes \citep{1985ApJ...298...18F}, or triggered by other mechanisms
such as shock compression or cloud-cloud colissions \citep[e.g.][]{
1980glcl.conf..301G, 1992ApJ...400..265M, 1996ASPC...92..241L, 1994ApJ...429..177H,
2001ApJ...560..592C}.

Yet another hypothesis is that star cluster formation is triggered by (major) 
gas-rich galaxy mergers \citep{1987nngp.proc...18S, 1992ApJ...384...50A},
which is naturally expected within the framework of hierarchical assembly. One
testable prediction of this scenario is the formation of young clusters in
interacting and merging galaxies, which has been observed and show remarkable
similarities with globulars in the Milky Way \citep[e.g.][]{1995AJ....109..960W, 
1996AJ....112..416H, 1999AJ....118..752Z, 1999AJ....118.1551W}. Moreover, models
that test this hypothesis yield GC populations consistent with various 
observables \citep[e.g.][]{2010ApJ...718.1266M, 2018MNRAS.480.2343C}, and
the recent numerical simulation of an isolated dwarf-dwarf merger executed at 
very high resolution (baryonic mass $m_b \sim$~4M\Sun; softening $\epsilon = 0.1$) 
produce promising star clusters that could be globular progenitors \citep{2019arXiv190509840L}.

More recent modelling efforts combine various aspects from the aforementioned
formation channels \citep[e.g.][]{2000ApJ...533..869C, 2002ApJ...567..853C, 2002MNRAS.333..383B,
2003egcs.conf..224G}

Previous studies
\citep{2005ApJ...623..650K}
\citep{2016ApJ...831..204R}
\citep{2017ApJ...834...69L}
\citep{2017MNRAS.465.3622R}
\citep{2018MNRAS.474.4232K}
\citep{2018MNRAS.475.4309P,2019MNRAS.486.3134K}

\todo[inline]{Paragraph: Paper outline}
We summarise the relevant characteristics of the Auriga simulations in
section~\ref{sec:auriga}, followed by a summary of the observations of the
Milky Way (MW) and Andromeda globular cluster system (GCS) in
section~\ref{sec:observations} that we use to compare our simulations to in
section~\ref{sec:results}. We discuss our findings in section~\ref{sec:discussion}
to come to our conclusions in section~\ref{sec:conclusions}.
